---
title: "Homework : Gamma Delta Hedging"
author: "Huyen Le"
date: "December 29, 2016"
output: 
  pdf_document:
    fig_caption: true
    number_sections: true
      
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = F, error = F, warning = F)
library('ggplot2')
library('tidyr')
library('moments')
load('Results for reports.Rdata')
```

#Assumptions

The result is based on the following assumption:

* Initial price $S_0=100$.

* Implied volatility is $\sigma_1 = 0.2$.

* Realized volatility (used for Monte Carlo simulation stock price path) is $\sigma_2 = 0.3$.

* Interest rate $r = 0.05$.

* The option to be hedged is one-year European call option with strike price $K = 100$ (Option 1).

* There are 240 working days in a year, each day has 6.5 trading hours.

The frequency used for re-hedging is:

* $f = 9630$: hedging every ten-minute period (*almost* continuous hedging)

* $f = 1560$: hedging every hour

* $f = 240$: hedging every day

* $f = 48$: hedging every week

Number of Monte Carlo simulations $N = 10000$.

#Delta Hedging and re-hedging frequency.

##Hedging with known realized volatility.

For simplicity, in this case, I assume that Option 1 is priced with real volatility $\sigma_2 = 0.3$ at time $t = 0$. Put this another way, at time $t =0$, $C_i = C_r$. Assuming that the initial position is going long for one unit of Option 1, the basic steps will be:

* Delta-hedge by shorting $\Delta_r$ units of underlying asset $S_t$, with *real* volatility of $\sigma_2$.

* Change $\Delta_r$ as $S_t$ and time to maturity (TTM) change over the life of the option.

* Finance (invest) any purchase (payment from going short) by borrowing (investing) at risk-free rate $r$ .
  
The below chart is the performance of the strategy in different frequency of re-hedge.

```{r plot1, fig.cap = "Delta Hedging by *real* volatility in different re-hedge frequency"}

PnL2.plot <- gather(data = PnL2, Frequency, PnL)

ggplot(PnL2.plot, aes(x = PnL, fill = Frequency)) +
  geom_histogram(binwidth = 0.01) +
  facet_wrap(~ Frequency, scales = "free_y") +
  ylab("Number of simulations")+
  xlim(-5,5)
```


**Conclusion**: in the case that we **DO** know the real volatility, **increase the re-hedging frequency will REDUCE the exposure to change in stock price**.

##Hedging with implied volatility (unknown real volatility)

In this case, I assume that Option 1 is priced with implied volatility $\sigma_1 = 0.2$ at time $t = 0$. Put this another way, at time $t =0$, $C_i \neq C_r$.

Assuming that the initial position is going long for one unit of Option 1, the strategy will include:

* Delta-hedge by shorting $\Delta_r$ units of underlying asset $S_t$, with $\Delta$ values are calculated based on *implied volatility* of $\sigma_1$

* Change $\Delta_r$ as $S_t$ and TTM change over the life of the option

* Finance(invest) any purchases (payments from going short) by borrowing (investing) at risk-free rate $r$ 

The chart below displays the performance of the strategy. It is noteworthy that the Profit and Loss in this case is positive only because real volatility $\sigma_2$ is greater than the implied one $\sigma_1$. That leads to the fact that we actually **under-hedged** and became less exposed to the underlying price. In the opposite case, we will suffer from loss due to over-hedging.

```{r plot2, fig.cap="Delta Hedging by implied volatility in different re-hedge frequency"}

ggplot(gather(data = PnL, Frequency, PnL), aes(x = PnL, fill = Frequency)) +
  geom_histogram(binwidth = 0.01) +
  facet_wrap(~ Frequency, scales = "free_y") +
  ylab("Number of simulations")
```


Also, following () is the comparison of PnL statistics in case of Hedging by real and implied volatility, respectively. The standard deviation of PnL was much higher and went **flat** even after increasing Delta rebalancing frequency for hedging using implied volatility.


```{r plot3, fig.cap = "Convergence of standard deviation of hedging error - Delta Hedging" ,fig.height = 3, fig.width = 5, echo = FALSE, fig.pos = "H"}
b<-data.frame(statistics2[2,], statistics[2,], c(n, n2, n3, n4))

names(b) <- c("real vol", "implied vol", "Frequency")


b <- gather(b, Type, Std, - Frequency)

ggplot(b, aes(x = Frequency, y = Std, col = factor(Type))) + 
    xlab("Frequency of Delta rebalancing")+
    theme(legend.title=element_blank())+
    geom_point()+
    geom_line()
```



**Conclusion**: As can be seen, in the case that we do **NOT** know the real volatility, **increase the re-hedging frequency will not help reduce the exposure to the changes in stock price**. 

# Gamma-Delta Hedging

From the above part, we can see that delta hedging, by doing continuously with real volatility, can reduce the exposure of the portfolio to the changes in underlying price.

In this part, I will implement Gamma-Delta hedging by establishing a delta and gamma neural portfolio using real volatility $\sigma_2 = 0.3$. To implement gamma-hedge, the second option is also European call with strike $K = 90$ (*Option2*). 

It is worth noting that my strategy is really simple, I only gamma hedge the portfolio **once** at the beginning and rebalance delta of the portfolio (which now contains Option 1 and Option 2) with different frequencies to see if I can reduce the hedging error for the portfolio. 

Assuming that the initial position is going long for one unit of Option 1, the basic steps will be:

* In order to make the portfolio gamma-neutral, we take position on Option 2 as follows: 

$$
Pos_{option 2} = -\frac{\Gamma_1}{\Gamma_2}
$$
By doing so, we make the portfolio gamma neutral at time $t =0$.

* Afterwards, take the position in underlying stock to delta hedge the option portfolio. In other words, position of underlying will be $Pos_{stock} = -\Delta_{portfolio}$.

* Change $Pos_{stock}$ over the life of the two options.

* Finance(invest) any purchase (payment from going short) by borrowing (investing) at risk-free rate $r$. 

In the following page is the figure for the performance of the Gamma-Delta neutral portfolio in this case and the line chart showing the convergence of hedging error's standard deviation for Delta and Delta-Gamma hedging strategies. It is confirmed in both that standard deviation of a Delta - Gamma hedged portfolio converges to zero more quickly than the Delta hedged portfolio


```{r plot4, fig.cap = "Delta - Gamma Hedging by known real volatility in different re-hedge frequency" , echo = FALSE}

PnL3.plot <- gather(data = PnL3, Frequency, PnL)


ggplot(PnL3.plot, aes(x = PnL, fill = Frequency)) +
  geom_histogram(binwidth = 0.01) +
  facet_wrap(~ Frequency, scales = "free_y") +
  ylab("Number of simulations")
```



```{r plot5, fig.cap = "Convergence of standard deviation of hedging error  - Delta and Delta-Gamma hedging", ,fig.height = 3, fig.width = 5, echo = FALSE}
c<-data.frame(statistics2[2,], statistics3[2,], c(n, n2, n3, n4))

names(c) <- c("Delta", "Delta-Gamma", "Frequency")


c <- gather(c, Type, Std, - Frequency)

ggplot(c, aes(x = Frequency, y = Std, col = factor(Type))) + 
    geom_point()+
    geom_line()+
    xlab("Frequency of Delta rebalancing")+
    theme(legend.title=element_blank())
```


**Conclusion**: the performance of Gamma-Delta hedging is much better (with smaller hedging error) for every frequency. Taking transaction cost into consideration, **we can reduce the trading frequency to achieve the same desired hedging error** by making our portfolio neutral in both gamma and delta. 

# Attachments

Being attached with this report is:

* code in ```C++``` to create the package in ```R``` for Monte Carlo Simulation and 

* code in ```R``` for calculation and visualization.



Thank you very much for reading (if you read this report until the end). 

Wish you and your family a Happy New Year!

___

\newpage

#Appendix

Here is the summary statistics for the strategies applied in this project


```{r}
knitr::kable(statistics, caption  = "Hedging with implied volatility statistics")
```

```{r}
knitr::kable(statistics2, caption  = "Hedging with known realized volatility statistics")
```

```{r}
knitr::kable(statistics3, caption  = "Gamma-Delta Hedging with known realized volatility statistics")
```




